---
title: "Regression"
output: html_document
date: 13/03
---


#Linear Regression
**how does mean of our outcome y vary with other variables**
**predict a quantitative outcome variable (y) on the basis of one or multiple predictor variables (x)**
**outcome variable - can have any one of an infinite number of possible values**
http://www.sthda.com/english/articles/40-regression-analysis/165-linear-regression-essentials-in-r
https://stat.ethz.ch/R-manual/R-devel/library/boot/html/melanoma.html

**simple linear regression (assignment 3)**
```{r setup, include=FALSE}
library(COUNT)
data(lbw)
class(lbw)
t(t(names(lbw)))
```


```{r, include=TRUE}
data(lbw)
View(lbw)
plot(lbw$lwt, lbw$bwt, main="Regression of infant birth weight and maternal weight at last menstrual period", xlab="Weight at last menstrual period", ylab="Birthweight (g)") +
abline(lm(bwt ~ lwt, data = lbw), col = "red")
#corr.test(lbw$bwt, lbw$lwt)
linmod<-lm(bwt ~ lwt, data = lbw)
summary(linmod)

#example 2
plot(lbw$ftv, lbw$bwt)
abline(lm(bwt ~ ftv, data = lbw))
linmod_a <- lm(bwt ~ ftv, data = lbw)
summary(linmod_a)
```





**1. build and summarize the linear model**
**predicting the thickness of the melanoma on the basis of age, sex and ulceration of melanoma - multiple (multivariable) linear regression**
**thickness is numerical - continuos**

```{r, include=TRUE}
data("melanoma")
linmod1<-lm(thickness ~ age + sex + ulcer, data = melanoma) #build a model
summary(linmod1) #summarize the model
attributes(linmod1)
attributes(linmod1$coefficients)
```

**2. Checking assumptions**
```{r, include=TRUE}
par(mfrow = c(2, 2)) #all4 diagnostic plots displayed together
plot(linmod1) # you get 4 diagnostic plots, not a plot of the model
```
-Residuals vs Fitted: check the linear relationship assumptions. ô°€ What to look for: horizontal line and no distinct patterns
-Normal Q-Q plot: check that the residuals are normally distributed.  What to look for: residuals points follow the straight dashed line
-What to look for: equally spread points
-Residuals vs Leverage: identify influential cases/extreme values. Extremes are labelled

**3.report as table**
```{r}
summary(linmod1)
library(kableExtra)
kable(summary(linmod1)$coef, digits = c(3, 3, 3, 4)) #report as table
#association between thickness and ulcer
```

####Practice 2
-Run a multivariable linear regression analysis to see if Species and Sepal Width are independently associated with Sepal Length. Make Versicolor the reference category

```{r, include=TRUE}
data(iris)
head(iris)
summary(iris)
class(iris$Species)
levels(iris$Species)
mlr <- lm(Sepal.Length ~ Sepal.Width + relevel(as.factor(Species), ref="versicolor"), data = iris)
summary(mlr)
```



#Logistic regression
**the outcome (dependent variable) has only a limited number of possible values** 
**Logistic regression is used when the response variable is categorical in nature.**
**predicting wether melanoma will be ulcerated (yes or no - categorical, event is binary)**
**status:1  died from melanoma, 2  alive and 3 died from causes unrelated to their melanoma.**
**1. reference by default is status 1 - died from melanoma**
```{r, include=TRUE}
table(melanoma$ulcer)
levels(as.factor(melanoma$status))
logmod1 <- glm(ulcer ~ age + as.factor(status) + as.factor(sex), family = binomial, data = melanoma)
summary(logmod1)
coef.table <- summary(logmod1)$coef 
kable(coef.table, digits=c(3, 3, 2, 4))
```


**2. change the reference category**
**now ref is status 3 i.e. died from unrel cuases**
```{r}
logmod2 <- glm(ulcer ~ age + relevel(as.factor(status), ref=3) + as.factor(sex), family = binomial, data = melanoma)
coef.table2 <- summary(logmod2)$coef 
coef.table2
kable(coef.table2, digits=c(3, 3, 2, 4))
```

**3. Coefficients are log-odds - exponentiate**
```{r}
## get odds ratios and 95% CI from logmod1 
exp(cbind(OR = coef(logmod1), confint(logmod1)))
logmod_fin<-exp(cbind(OR = coef(logmod1), confint(logmod1))) 
kable(logmod_fin)
```












